Capon Beamforming Applied on Moving Objects - An Analysis of Local Lateral Shift-Invariance

### 2013-11-21 HRUIofMovingObjects/jrnl_article.tex
Moving object article: Done some writing. Added new figures

\author{
   Jon Petter \AA{}sen, \IEEEmembership{Student Member,~IEEE,}, Andreas Austeng, \IEEEmembership{Member,~IEEE,} and Sverre Holm, \IEEEmembership{Senior Member,~IEEE,}% <-this % stops a space

@@

\begin{abstract}
%\boldmath
If an ultrasound imaging system provides a presentation of a moving object witch is not invariant to local spatial shifts the system is said to be local shift-variant. This can for instance happen if the axial and lateral sampling is insufficient. The Capon beamformer has been shown to provide increased lateral resolution in ultrasound images. Increased lateral resolution should demand denser lateral sampling. In previous literature on Capon beamforming for medical ultrasound imaging however only single frame scenarios have been simulated, and temporal behaviour and effects caused by the increased resolution and lack of oversampling has therefore been neglected.

@@

\IEEEPARstart{W}{ith} the introduction of a fast implementation of the Capon adaptive beamformer, a stream of images can now be processed at interactive frame rates \cite{Asen}. It is therefore an interesting question how a moving object appears after Capon beamforming has been applied. Earlier work on Capon beamforming for medical ultrasound imaging has investigated the method's capabilities in single frames only, and temporal behaviour and effects has therefore been neglected.

An important property of an ultrasound imaging system is its local shift-invariance. Local shift-invariance means that an object’s appearance should be preserved when the object, or the ultrasound probe, is moved slightly. This invariance for local movements is what makes ultrasound imaging feasible in general, however it is specially important for quantitative techniques that correlate time-varying features, like pulsed doppler and flow imaging, and speckle tracking of tissue and blood. In this paper we analyze the shift-invariance property of the Capon beamformer by studying the beamformer output when applied on simulated and \textit{in vitro} images of moving objects. We also show some examples of how the shift-invariance and resolution is improved in \textit{in-vivo} images.

Capon beamforming has been shown to provide increased lateral resolution and contrast in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. The method accomplish this by adjusting the array weights based on element-to-element covariance. Hence, the method yields data dependent weights. Capon beamforming is however also known for its suboptimal performance if the sample size available for covariance estimation is limited \cite{Mestre2006}, if correlated signals are present \cite{Widrow1982}, and if the assumed steering vector does not match the signal propagation vector \cite{Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets (SVUS) \cite{Lorenz2005, Rubsamen2013}. The first two, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics \cite{Synnevag2007a}. SVUS has not been applied, even though it is well known that the Capon beamformer exhibits severe signal-nulling if the signal is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the method obtains its super resolution capabilities.

In \cite{Cox1973}, Cox derives formulas for the resolution of the Capon beamformer (called the $\vec{k}_3$ processor), and presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. Cox states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the steered response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to three times the system resolution. Impinging on the array is a monochromatic plane wave with a propagation vector orthogonal to the array surface with no spatially white noise present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$) is the identity. The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) has also been added to the Capon beamformer output. A diagonal loading of $1/100$ corresponds to an signal-to-noise ratio of 40 dB before array gain is applied \comment{Is this correct?}. The figure shows that when the DAS beamformer is used, only a 3.92 dB scalloping loss \cite{Harris1978} should be observed if the source of the plane wave starts moving parallel to the array. However if the Capon beamformer is used with the beam density sufficient for DAS the scalloping loss could be as large as $31$ dB.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img steerd_respons_capon_das_resolution.eps}
}
\caption{Comparison of drop in output power in between beams (half the system resolution) for delay-and-sum (DAS) and Capon beamforming ($M=64$, $L=32$, $d=0.01$). A diagonal loading of $1/100$ corresponds to an effective signal-to-noise ratio after array gain of 40 dB.}
\label{fig:das_capon_beams}
\end{figure}

In previous literature on Capon beamforming for medical ultrasound imaging (with single frame simulations) this effect has been passed over in silence by either oversampling on transmit, or by carefully position point scatterers exactly on transmit-receive beam pairs, or by ignorance. In our first work where Capon beamforming was applied on a loop of images containing moving objects \cite{Asen2012, Asen} we commented on the effect, but no solution was presented other than oversampling on transmit when simulating.

@@

In the next section we give an introduction to Capon beamforming for medical ultrasound imaging together with a short introduction to lateral sampling and shift invariance. In Section \ref{sec:capon_LLSI} the local lateral shift-invariance of the Capon beamformer is investigated, before the different methods are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res} and Section \ref{sec:dis} we present and discuss the results of applying the phase rotation method on \textit{in-vitro} data . Finally we give our conclusion in Section \ref{sec:con}.

@@

\subsection{Capon Beamforming}
The Capon beamformer produce a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criteria. The minimization problem is as follow:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is a sample covariance matrix. Hence, the resulting weight vector will minimizes the output power while maintaining unit gain in the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

@@

Note that one matrix has to be constructed and inverted for each data vector $\vec{x}[n]$ received by the system. This is indeed computational demanding, but following the innovation in GPU computing, and by applying a novel method in order to reduce the size of the matrix inversion problem \cite{Nilsen2009}, real time processing is now possible \cite{Asen}. From (\ref{eq:capon_weights} it is clear that complex data will result in complex weights. The weights can therefore have an asymmetric K-space response which basically means that zeros does not have to be symmetric and that the beam can be micro-steered by phase rotation.

@@

The lateral sampling required in an ultrasound image is given by the combined aperture size on transmit ($D_{tx}$) and receive ($D_{rx}$), and the center frequency ($f_c = c/\lambda$), where $c$ is the speed of sound in the body and $\lambda_c$ is the wavelength of the center frequency. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{D_{tx} + D_{rx}}. \label{eq:resolution}
\end{align}
If this requirement is not fulfilled the scalloping loss could become visible in the image and the system will no longer be lateral shift-invariant. In an ultrasound image with 256 gray levels and 50 dB dynamic range a 1 dB loss corresponds to 5 gray levels. This is approximately equal to the visibility threshold (Weber fraction of 2\%) for a grayscale image, and a loss larger than 1dB could therefore end up being visible to the observer. In our analysis a system would, based on this reasoning, be regarded as lateral shift-invariant if the loss is smaller than 1 dB. However, depending on light conditions and e.g. the selected gray level mapping, a larger loss might be acceptable.

@@

Fig.\,\ref{fig:das_shiftinvariant} presents a LSV-plot for an imaging system using DAS beamforming with a beam spacing equal to (\ref{eq:resolution}). The red dotted vertical lines indicates the position of tx-rx beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} \comment{add table with simulation settings or just insert in text?}. We see that the contours are approximately diagonal and therefore we can conclude that the system is locally lateral shift-invariant. The P-MAE and A-MAE are also minimal.

@@

\subsection{Oversampling on Transmit}
The straight forward way of obtaining sufficient sampling in an active system is by increasing the number of transmit beams (tx-beams) until (\ref{eq:resolution}) is fulfilled or the LSV-plot has a uniform level along the diagonals. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the lateral sampling on transmit by a factor of four and ten respectively. As expected the scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

From Fig.\,\ref{fig:capon_oversampling_b} we see that more than ten times oversampling has to be applied to achieve local lateral shift-invariance when a moving point scatterer is imaged without noise and with the same parameters as in Fig\,\ref{fig:capon}. %$1/100$ in diagonal loading. 
Note that a diagonal loading factor of $1/100$ will introduce a deterministic white noise component in the sample covariance matrix $40$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 40 dB.

Even though increasing the number of transmits improves the lateral shift-invariance it will also reduce the imaging frame rate with a factor equal to the oversampling factor. The processing complexity will also increase equally. For these two reasons, oversampling on transmit is not a feasible option for real time ultrasound imaging.

@@

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rates in e.g. cardiac ultrasound imaging. However, PRB also introduce motion artifacts in the image which to some degree can be compensated for \cite{Hergum2007} \comment{cite Bastien}. In (\comment{Cite paper where they use STB + Capon if its out}) it is shown that synthetic transmit beamforming (STB) \cite{Hergum2007} works with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In this discussion we consider PRB as being equal to oversampling on transmit when it comes to image quality.

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams while the number of tx-beams are reduced. When using PRB for oversampling we would like to keep the number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand more rx-beams will increase the number of covariance matrices, and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with a factor equal to the number of parallel receive lines. Since almost 10x oversampling is required in order to achieve lateral shift invariant imaging of a moving point scatterer, non of the implementations described in \cite{Asen} would be real time any more.

One possible solution is to estimate the covariance matrix from multiple beams \cite{Jensen2012}, generating one covariance matrix per range.  Tailoring the method of Jensen and Austeng to cardiac ultrasound imaging it makes more sense to construct one covariance matrix per tx-beam using multiple parallel rx-beams. This solution will not increase the number of matrix inversions, but the time taken to construct the matrices will increase. We will not investigate this method further in this paper. \comment{Kanskje litt dumt å nevne det da...?}

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by only using phase rotation. This is also possible in broadband applications as long as the phase rotation is limited to a fraction of the wave length. It is therefore possible to make a sweep between the current beam and its neighbours using only phase rotation. Phase rotation is what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance cyst edges.

For a uniform linear array located along the x-axis the steering vector $\vec{a}$ should be calculated as 
\begin{align}
\vec{a} = 
\begin{bmatrix}
e^{-j\frac{2\pi}{\lambda_c}x_0\sin(\theta)} \\
e^{-j\frac{2\pi}{\lambda_c}x_1\sin(\theta)} \\
\vdots \\ 
e^{-j\frac{2\pi}{\lambda_c}x_{L-1}\sin(\theta)}
\end{bmatrix}
\end{align}
where $x_i$ is the element position and $\theta$ is swiped from $-\Delta/2$ too $\Delta/2$.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{\includegraphics[width=0.3\linewidth]{\img phantom1_standard_processing.png}\label{fig:phantom_das}}
\hfill{}
\subfloat[]{\includegraphics[width=0.3\linewidth]{\img LSV_capon.png}\label{fig:phantom_capon_LSV}}
\hfill{}
\subfloat[]{\includegraphics[width=0.3\linewidth]{\img LSV_capon.png}\label{fig:phantom_capon_LSV_8PR}}
}
\caption{\comment{Not final figure. Need new data}. LSV-plots of in-vitro data. a) Image of the data. b) Normal capon. c) Capon with oversampling using 8PR.}
\label{fig:das}
\end{figure*}

@@

The authors would like to thank...

>>>>

\author{
   Jon Petter \AA{}sen, \IEEEmembership{Student Member,~IEEE,}, Andreas Austeng, \IEEEmembership{Member,~IEEE,} and Sverre Holm, \IEEEmembership{Senior Member,~IEEE}% <-this % stops a space

@@

\begin{abstract}
%\boldmath
If an ultrasound imaging system provides a presentation of a moving object which is sensitive to local spatial shifts, the system is said to be local shift-invariant. This can for instance happen if the axial and lateral sampling is insufficient. The Capon beamformer has been shown to provide increased lateral resolution in ultrasound images. Increased lateral resolution should demand denser lateral sampling. In previous literature on Capon beamforming for medical ultrasound imaging however only single frame scenarios have been simulated, and temporal behaviour and effects caused by the increased resolution and lack of oversampling have therefore been neglected.

@@

\IEEEPARstart{I}{n} the last decade several authors have applied the Capon beamformer for ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The adaptive nature of the method has been shown to greatly improve the lateral resolution in ultrasound images. In recent work we introduced a fast implementation of the Capon beamformer, capable of processing a stream of images at interactive frame rates \cite{Asen}. It is therefore an interesting question how a moving object appears after Capon beamforming has been applied. Earlier work on Capon beamforming for medical ultrasound imaging has investigated the method's capabilities in single frames only, and temporal behaviour and effects has therefore been neglected.

An important property of an ultrasound imaging system is its local shift-invariance. Local shift-invariance means that an object's appearance should be preserved when the object, or the ultrasound probe, is moved slightly. This invariance for local movements is what makes ultrasound imaging feasible in general, but it is specially important for quantitative techniques that correlate time-varying features, like pulsed doppler and flow imaging, and speckle tracking of tissue and blood. In this paper we analyze the lateral shift-invariance property of the Capon beamformer by studying the beamformer output when applied on simulated and \textit{in vitro} images of moving objects. We also show some \textit{in-vivo} examples.

As mentioned, Capon beamforming has been shown to provide improved lateral resolution and edge definitions in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. The method accomplishes this by adjusting the array weights based on element-to-element covariance. Hence, the method yields data dependent weights. Capon beamforming is however also known for its suboptimal performance if the sample size available for covariance estimation is limited \cite{Mestre2006}, if correlated signals are present \cite{Widrow1982}, and if the assumed steering vector does not match the signal propagation vector \cite{Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets (SVUS) \cite{Lorenz2005, Rubsamen2013}. The first two, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics \cite{Synnevag2007a}. SVUS reduces the signal-nulling effect of the Capon beamformer which appears if the signal is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the method obtains its super resolution capabilities, but the signal can be lost in the process. In this paper we make use of a steering vector set to reduce this self-nulling effect.

In \cite{Cox1973}, Cox derives formulas for the resolution of the Capon beamformer, and presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. Cox states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the steered response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to three times the system resolution. Impinging on the array is a monochromatic plane wave with a propagation vector orthogonal to the array surface with no spatially white noise present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) has also been added to the Capon beamformer output. A diagonal loading of $1/100$ corresponds to an signal-to-noise ratio of 20 dB before array gain is applied. The figure shows that when the DAS beamformer is used with a beamspacing equal to the system resolution, only a 3.92 dB scalloping loss \cite{Harris1978} should be observed if the source of the plane wave starts moving parallel to the array. However if the Capon beamformer is used with the same beam density the scalloping loss could be as large as $31$ dB.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img steerd_respons_capon_das_resolution.eps}
}
\caption{Comparison of drop in output power in between receive beams. Half the system resolution is plotted to show where two beams with a spacing equal to the system resolution would intersect. The steered response of both delay-and-sum (DAS) and Capon beamforming ($M=64$, $L=32$, $d=0.01$) is plotted. A monochromatic plane wave is located at zero degrees.}
\label{fig:das_capon_beams}
\end{figure}

In previous literature on Capon beamforming \comment{cite earlier work} for medical ultrasound imaging (with single frame simulations) this effect has been passed over in silence by either oversampling on transmit, or by carefully position point scatterers exactly on transmit-receive beam pairs, or by ignorance. In our first work where Capon beamforming was applied on a loop of images containing moving objects \cite{Asen2012, Asen} we commented on the effect, but no solution was presented other than oversampling on transmit when simulating.

@@

In the next section we give an introduction to Capon beamforming for medical ultrasound imaging together with a short introduction to lateral sampling and shift invariance. In Section \ref{sec:capon_LLSI} the local lateral shift-invariance of the Capon beamformer is investigated, before the different methods are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res} and Section \ref{sec:dis} we present and discuss the results of applying the most promising method on \textit{in-vitro} and \textit{in-vivo} data . Finally we give our conclusion in Section \ref{sec:con}.

@@

\subsection{Capon Beamforming}
The Capon beamformer produce a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criteria. The minimization problem is as follow \cite{Capon1969}:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is a sample covariance matrix. Hence, the resulting weight vector will minimizes the output power while maintaining unit gain in the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

@@

Note that one matrix has to be constructed and inverted for each data vector $\vec{x}[n]$ received by the system. This is indeed computational demanding, but following the innovation in GPU computing, and by applying a novel method in order to reduce the size of the matrix inversion problem \cite{Nilsen2009}, real time processing is now possible \cite{Asen}. From (\ref{eq:capon_weights}) it is clear that complex data will result in complex weights. The weights can therefore have an asymmetric K-space response which basically means that zeros does not have to be symmetric around the origin and that the beam can be micro-steered by phase rotation.

@@

The lateral sampling required in an ultrasound image is given by the combined aperture size on transmit ($D_{tx}$) and receive ($D_{rx}$), and the center frequency ($f_c = c/\lambda$), where $c$ is the speed of sound in the body and $\lambda_c$ is the wavelength of the center frequency. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{D_{tx} + D_{rx}}. \label{eq:resolution}
\end{align}
If this requirement is not fulfilled the scalloping loss could become visible in the image and the system will no longer be lateral shift-invariant. In an ultrasound image with 256 gray levels and 50 dB dynamic range a 1 dB loss corresponds to 5 gray levels. This is approximately equal to the visibility threshold (Weber fraction of 2\%) for a grayscale image, and a loss larger than 1dB could therefore end up being visible to the observer. In our analysis a system would, based on this reasoning, be regarded as locally lateral shift-invariant if the beam-too-beam loss is smaller than 1 dB. However, depending on light conditions and e.g. the selected gray level mapping, a larger loss might be acceptable.

@@

Fig.\,\ref{fig:das_shiftinvariant} presents a LSV-plot for an imaging system using DAS beamforming with a beam spacing equal to (\ref{eq:resolution}). The red dotted vertical lines indicates the position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} \comment{add table with simulation settings or just insert in text?}. We see that the contours are approximately diagonal and therefore we can conclude that the system is locally lateral shift-invariant. The P-MAE and A-MAE are also minimal.

@@

\subsection{Oversampling on Transmit}
The straight forward way of obtaining sufficient sampling in an active system is by increasing the number of transmit beams (tx-beams) until (\ref{eq:resolution}) is fulfilled or the LSV-plot has an uniform level along the diagonals. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the lateral sampling on transmit by a factor of four and ten respectively. As expected the scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

From Fig.\,\ref{fig:capon_oversampling_b} we see that more than ten times oversampling has to be applied to achieve local lateral shift-invariance when a moving point scatterer is imaged without noise and with the same parameters as in Fig\,\ref{fig:capon}. %$1/100$ in diagonal loading. 
Note that a diagonal loading factor of $1/100$ will introduce a deterministic white noise component in the sample covariance matrix $40$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 40 dB. \comment{correct?}

Even though increasing the number of transmits improves the lateral shift-invariance it will also reduce the imaging frame rate with a factor equal to the oversampling factor. The processing complexity will also increase equally. For these two reasons, oversampling on transmit is not a feasible option for real time ultrasound imaging, and we will not discuss this method further.

@@

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rates in e.g. cardiac ultrasound imaging. However, PRB also introduce motion artifacts in the image which to some degree can be compensated for \cite{Hergum2007} \comment{cite Bastien}. In (\comment{Cite paper where they use STB + Capon if its out}) it is shown that synthetic transmit beamforming (STB) \cite{Hergum2007} works with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In the following discussion we consider PRB as being equal to oversampling on transmit when it comes to image quality.

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams while the number of tx-beams are reduced. When using PRB for oversampling we would like to keep the number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand more rx-beams will increase the number of covariance matrices, and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since almost 10x oversampling is required in order to achieve lateral shift invariant imaging of a moving point scatterer, non of the implementations described in \cite{Asen} would be real time any more.

One way of reducing the number of matrix inversions is to estimate the covariance matrix from multiple beams \cite{Jensen2012}, generating one covariance matrix per range.  Tailoring the method of Jensen and Austeng to cardiac ultrasound imaging it makes more sense, due to motion, to construct one covariance matrix per tx-beam using multiple parallel rx-beams. With this solution the number of matrix inversions will not changes, but the time taken to construct the matrices will increase. We will not investigate this method further in this paper. \comment{Kanskje litt dumt å nevne det da...?} \comment{move to discussion}

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by only using phase rotation. This is also possible in broadband applications as long as the phase rotation is limited to a fraction of the wave length \comment{cite fine-coarse beamformer}. It is therefore possible to make a sweep between the current beam and its nearest neighbours using only phase rotation. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

For an uniform linear array located along the x-axis the steering vector $\vec{a}$ should be calculated as 
\begin{align}
\vec{a} = 
\begin{bmatrix}
e^{-j\frac{2\pi}{\lambda_c}x_0\sin(\theta)} \\
e^{-j\frac{2\pi}{\lambda_c}x_1\sin(\theta)} \\
\vdots \\ 
e^{-j\frac{2\pi}{\lambda_c}x_{L-1}\sin(\theta)}
\end{bmatrix}
\end{align}
where $x_i$ is the element position and $\theta$ is swiped from $-\Delta/2$ too $\Delta/2$ where $\Delta$ is as defined in (\ref{eq:resolution}). Note that the steering vector is calculated for a $L$ element array centered around the origin, even though the full array has $M$ elements. It is important that this synthetic array has the same element pitch as the full array.

@@

\begin{figure*}[!t]
\centerline{
\begin{tabular}{cc}
\subfloat[]{\includegraphics[width=0.50\linewidth]{\img LSV_image.eps}\label{fig:phantom_das}}
\hfill{}
\subfloat[]{\includegraphics[width=0.50\linewidth]{\img LSV_das.eps}\label{fig:phantom_capon_LSV}}\\
\hfill{}
\subfloat[]{\includegraphics[width=0.50\linewidth]{\img LSV_capon.eps}\label{fig:phantom_capon_LSV}}
\hfill{}
\subfloat[]{\includegraphics[width=0.50\linewidth]{\img LSV_capon_10pr.eps}\label{fig:phantom_capon_LSV_8PR}}
\end{tabular}
}
\caption{LSV-plots generated from \textit{in-vitro} phantom data. The probe was manually dragged across the phantom. a) First frame. b) DAS c) Normal capon. d) Capon with oversampling using 10 times phase rotation.}
\label{fig:das}
\end{figure*}

@@

The authors would like to thank Anders Sornes for acquiring data.
